Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs

### 2014-07-11 Referee2_Response.tex

{\large Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs}
\medskip

{\footnotesize by D. Love and G. Bayraksan}
\end{center}

\bigskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noi 
We thank the referee for the helpful comments and suggestions. 
We incorporated all of the recommended changes and believe the paper is improved as a result. 
Below, we go through in detail these changes. 
\medskip

@@

%%%%%%%%%%%%%%%%%
%% Need to be careful with the minor changes. Will these be affected?
%% Need to CHECK AT THE VERY END... FOR ANY CHANGES on TOP OF these...
%%    - GB
%%%%%%%%%%%%%%%%%

\noi  
{\large \bf General Comments:}
\medskip

\begin{itemize}
\item[1.] \textit{The main contributions of the paper need to be stated more clearly and succinctly.}
\end{itemize}

\noindent 
You are right. 
We changed the introduction to clearly state the main contributions of the paper.  
Please see page four of the revised manuscript. 
We also clarified the motivation for the work, included new contributions based on the study computational efficiency added to the paper (contribution (iv)), and provided a comparison with the proceedings version of the paper (based on the editor-in-chief's suggestion).
%We believe the presentation of the paper significantly improved.  
%Thank you.
\medskip

@@

{\it Note:} While making the changes, we realized that `$\alpha$' was used in the current equation (3) and the statement of Theorem 2 as well as the ($1-\alpha$)-level confidence intervals.  
Therefore, we changed `$\alpha$' in equation (3) and Theorem 2 to `$\beta_{\bar{\gamma}}$'. 
\end{itemize}

\noi
With these changes, the OBM review became more streamlined, focusing on the main results used later in the paper. 
Thank you for this suggestion.
\medskip

\begin{itemize}
\item[3.] \textit{The English needs to be tightened up somewhat.}
\end{itemize}

\noindent  
We went through the paper and tightened up the prose. 
We also incorporated all the changes you suggested below and believe the exposition of the paper is improved as a result.
Thank you for your careful review that significantly improved the presentation of the paper. 
\medskip

@@

\begin{itemize}
\item[] \textit{2, 5: ``never before been studied . . . context.''}
\end{itemize}

\noi
Changed.
\medskip

\begin{itemize}
\item[] \textit{2, 7: By ``same benefits'', you mean moderate bias and comparatively low variance, right?}
\end{itemize}

\noi
Yes; we reworded this sentence. 
xxx We also went through the paper and made the context clear whenever a phrase like `same benefits', `desired results', etc.\ was used.  
\medskip

@@

\noi
Done.
\medskip

\begin{itemize}
\item[] \textit{2, 24: ``under which the resulting gap estimators exhibit reduced variance when compared to their counterparts obtained via nonoverlapping batches.''}
\end{itemize}

\noi
Done.
\medskip

\noi
{\it Note:}  We moved this sentence from the paragraph on page 2 to later in the introduction to clearly state the contributions of the paper. 
It now appears as contribution (ii) on page 4 of the revised manuscript.
\medskip

@@

\begin{itemize}
\item[] \textit{2, 24 (and elsewhere): Note: I would be careful about the use of the term ``variance reduction'', which usually refers to the class of techniques such as common random numbers, antithetics, importance sampling, etc. 
I view what you're doing in this paper as simply proposing a new estimator whose variance is lower than that of a competing estimator.}
\end{itemize}

\noi
xxx  You are right. 
Thanks for pointing this out. We corrected the wording throughout the paper. 
\medskip

\begin{itemize}
\item[] \textit{2, 27: ``and that the interval \ldots valid. 
We also present''}
\end{itemize}

\noi
We made the change to the first sentence above. 
The first sentence now appears as part of contribution (ii) and the second sentence as part of contribution (iii) on page 4 of the revised manuscript. 
We decided not to include `also' in the second sentence as it starts a new set of contributions. 
\medskip

@@

\begin{itemize}
\item[] \textit{2, 28: ``Our experiments indicate that the use of overlapping estimators results in gap estimators achieving lower variance than their nonoverlapping counterparts, along with comparable bias and interval coverage --- for both small- and large-sample cases.}
\end{itemize}

\noi
Done.
\medskip

\noi
{\it Note:} We moved this sentence from the paragraph on page 2 to later in the introduction to clearly state the contributions of the paper.
This sentence now appears as part of contribution (iii) on page 4 of the revised manuscript.
\medskip

@@

\noi
Deleted.
\medskip

\begin{itemize}
\item[] \textit{4, 17: ``interest is that of estimating''}
\end{itemize}

\noi
Done.  
\medskip

\noi
{\it Note:}
 As a result of the revised introduction---aimed to clearly state the contributions of the paper---we moved the technical discussion on the differences between using overlapping batches in simulation output analysis and assessing solution quality in stochastic optimization to \S 3 (Overlapping Multiple Replications Procedure). 
Consequently, this sentence now appears in \S 3. 
\medskip

@@

\begin{itemize}
\item[] \textit{4, 20--27: Even though I believe everything, these lines are a bit awkward and confusing.}
\end{itemize}

\noi
We removed the comment on the central limit theorem on the minimized sample means as it is not relevant to the paper. 
We also reworded these sentences to improve clarity. 
\medskip

\noi
{\it Note:}
As mentioned in the previous comment, this section now appears in  \S 3 (Overlapping Multiple Replications Procedure). 
\medskip

@@

\begin{itemize}
\item[] \textit{4, 31: What are the desired statistical properties?}
\end{itemize}

\noi
Thanks for pointing out the ambiguity here. 
We reworded this sentence, divided it into two sentences to improve readability, and added further explanations to remove any ambiguities. 
\medskip

\noi
{\it Note:}
This sentence now appears in \S 3 (Overlapping Multiple Replications Procedure). 
\medskip

### 2014-07-11 AE_Reponse.tex
Finishing up EinC, AE, Ref1 responses
and a final pass through Ref2 responses and comments -- Still not done.

We thank the Associate Editor for the helpful comments and suggestions. 
We incorporated virtually all of the recommended changes and believe the paper is improved as a result. 
Below, we go through in detail the suggestions of the Associate Editor.
\medskip

@@

\noi xxx 
We revised the introduction to clearly reflect the contributions of the paper and the motivation for the work, based on your and the referees' comments and the new material like computational efficiency added to the paper.
\medskip

\begin{itemize}
\item[] \textit{%\ldots
The experiments show that the confidence interval average widths and observed coverage probabilities are the same with overlapping batch means; the only difference is in decreased variability of the confidence interval widths. 
That is, the variance of the variance estimator is reduced. 
This second-order benefit should be weighed against the increased computation cost (not to mention complexity and restricted applicability) of the overlapping batch means method.}

\textit{The issue of computational efficiency is not addressed  in the paper. 
In this setting the efficiency is often defined as the reciprocal of the computation time multiplied by the variance. 
If this quantity were compared (overlapping batch means versus batch means) in the numerical experiments, I am curious if there are any examples where the efficiency is improved.}
\end{itemize}

\noi 
Thank you for this suggestion. 
It is indeed important to study computational efficiency in this context. 
We created a new section, Section 5.4 (Computational Efficiency), to study the computational efficiency of overlapping estimators in the stochastic optimization setting. 
When an algorithm has warm-starting capability, as the degree of overlapping increases,  solution time per batch could decrease; as a result, computational efficiency gains could be observed.  
We discussed efficiency for a given functional form of the solution time per batch. 
Depending on the effectiveness of warm-starting, maximizing computational efficiency provides guidelines on selecting the degree of overlapping. 
We illustrated this method on the newsvendor problem using insertion sort. 
Our experiments show that (i) the functional form used provides a good representation, (ii) increased computational efficiency over the nonoverlapping counterpart is indeed observed over a range of overlapping levels, and (iii) this analysis provides a way to select the degree of overlapping. 
We discussed guidelines on selecting the degree of overlapping (for NVD or for a given generic problem) in Section 5.4 (Computational Efficiency). 
\medskip

In addition, based on the comments of Referee 1, we also studied the efficiency of overlapping compared to increasing the number of batches in MRP to obtain (nonoverlapping) estimators with lower variances. 
Our results indicate that when warm-starting is effective, overlapping is computationally more efficient than increasing the number of batches of MRP.
We also illustrated this result on the newsvendor problem. 
We believe the paper has improved as a result of the study of computational efficiency. 
Thank you. 
\medskip

@@

\noi xxx 
Based on your other comments and suggestions of the referees, we provided information on computational efficiency gains of overlapping in the context of assessing solution quality in stochastic programming. 
These are detailed in Section 5.4 (Computational Efficiency). 
(Please see our response to the previous comment.)
We also changed Section 1 (Introduction) and Section 6 (Conclusions) to clearly motivate the study and usefulness of the results.

%xxx  Is this really a new comment?  Looks like just continuation of the prev. comment. So... combine it with previous comment???

\medskip

>>>>

\noi
We thank the Associate Editor for the helpful comments and suggestions. 
We incorporated all of the recommended changes and believe the paper is improved as a result. 
Below, we go through in detail the suggestions of the Associate Editor.
\medskip

@@

\noi
We revised the introduction to clearly reflect the contributions of the paper and the motivation for the work. 
Please see page four of the revised manuscript. 
This section also includes contributions based on the new material on computational efficiency added to the paper (contribution (iv)) and a comparison with the proceedings version of the paper (based on the editor-in-chief's suggestion).\medskip

\begin{itemize}
\item[] 
%\textit{
%The experiments show that the confidence interval average widths and observed coverage probabilities are the same with overlapping batch means; the only difference is in decreased variability of the confidence interval widths. 
%That is, the variance of the variance estimator is reduced. 
%This second-order benefit should be weighed against the increased computation cost (not to mention complexity and restricted applicability) of the overlapping batch means method.
%}
\textit{\ldots The issue of computational efficiency is not addressed  in the paper. 
In this setting the efficiency is often defined as the reciprocal of the computation time multiplied by the variance. 
If this quantity were compared (overlapping batch means versus batch means) in the numerical experiments, I am curious if there are any examples where the efficiency is improved.}
\end{itemize}

\noi 
Thank you for this suggestion. 
It is indeed important to study computational efficiency in this context. 
We created a new section, \S 5.4 (Computational Efficiency), to study the computational efficiency of overlapping estimators in the stochastic optimization setting. 
%When an algorithm has warm-starting capability, as the degree of overlapping increases,  solution time per batch could decrease; as a result, computational efficiency gains could be observed.  
We discussed efficiency for a given functional form of the solution time per batch. 
%Depending on the effectiveness of warm-starting, maximizing computational efficiency provides guidelines on selecting the degree of overlapping. 
%\medskip 
In addition, based on the comments of Referee 1, we studied efficiency of overlapping estimators compared to increasing the number of batches in MRP to obtain (nonoverlapping) estimators with equivalently lower variances. 
Our results indicate that when warm-starting is effective, overlapping is computationally more efficient.
Even in cases where the effectiveness of warm-starting is quite low, computational efficiency can be observed with a low degree of overlap (e.g., when approximately 10\% of observations are overlapped).\bigskip

\noi 
We illustrated these results on the newsvendor problem using insertion sort. 
Our experiments show that (i) the functional form used provides a good representation, (ii) increased computational efficiency over the nonoverlapping counterpart is indeed observed over a range of overlapping levels, and (iii) this analysis provides a way to select the degree of overlapping by maximizing efficiency.\bigskip

\noi
Finally, we provided guidelines on selecting the degree of overlap at the end of \S 5.4 (Computational Efficiency) based on our study. 
We believe the paper has improved as a result of the study of computational efficiency. 
Thank you for this suggestion. 
\medskip

@@

\noi 
We established improved computational efficiency on the newsvendor problem using insertion sort.
We also (i) changed Section 1 (Introduction) to clearly motivate the study and argue usefulness of the results, paying specific attention to efficiency gains; (ii) updated Section 6 (Conclusions) to summarize these new results, again paying attention to computational efficiency; (iii) provided a study of computational efficiency gains of overlapping in the context of assessing solution quality in stochastic programming in the new \S 5.4 (Computational Efficiency); and finally (iv) updated the abstract.  
%(Please see our response to the previous comment.)
\medskip

### 2014-07-11 EinC_Reponse.tex
Finishing up EinC, AE, Ref1 responses
and a final pass through Ref2 responses and comments -- Still not done.

@@

We thank the Editor-in-Chief for the helpful comments and suggestions. 
We incorporated all the recommended changes and believe the paper is improved as a result. 
Below, we go through in detail the suggestions of the Editor-in-Chief.\medskip

@@

\noindent xxx \ \ 
 We incorporated virtually all the recommended changes suggested by you, the associate editor, and the referees. 
We very much appreciate everyone's time and constructive comments. 
We believe the paper is significantly improved and the wording of the paper, compared to the Winter Simulation Conference paper, is changed as a result of the revision.
We also went through the paper and tried to change the text to reduce word for word copying. 
  \medskip

\begin{itemize}
\item[]\textit{Even more important, in clarifying your contribution (as already asked by the reviewers), you should also shortly explain what new research results are presented in the  TOMACS paper in comparison to the WSC paper}
\end{itemize}

\noindent
xxx Do xxx  We added a paragraph (??) in the introduction, detailing what has been done over \ldots

>>>>

\newcommand{\noi}{\noindent}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noi
We thank the Editor-in-Chief for the helpful comments and suggestions. 
We incorporated all the recommended changes and believe the paper is improved as a result. 
Below, we go through in detail the suggestions of the Editor-in-Chief.\medskip

@@

\noindent  
We incorporated all the recommended changes suggested by you, the associate editor, and the referees and believe the paper is significantly improved as a result.
We very much appreciate everyone's time and constructive comments. 
As a result of the revision, the wording and content of the paper, compared to the Winter Simulation Conference Proceedings version, has partly changed.
We also edited the paper to improve exposition and tried to cut back word for word copying during this process.  
\medskip

\begin{itemize}
\item[]\textit{Even more important, in clarifying your contribution (as already asked by the reviewers), you should also shortly explain what new research results are presented in the  TOMACS paper in comparison to the WSC paper (please note TOMACS expects about 30\% new research).}
\end{itemize}

\noindent
We added a paragraph in Section 1 (Introduction) detailing what has been done over the WSC proceedings paper. 
Please see the first paragraph on page 5 of the revised manuscript.\bigskip

\noi 
Specifically, we now show (i) almost sure consistency of the variance estimator (earlier this was mean-square consistency), (ii) asymptotic validity of the resulting confidence intervals, and (iii) that overlapping MRP variance estimators have asymptotically lower variances compared to (nonoverlapping) MRP variance estimator. 
These new theoretical results are due to a stronger assumption made in the paper submitted to TOMACS, which was not present in the WSC version. 
We also present a number of new computational results. 
For instance, we (iv) provide computations on two additional test problems, including a large-scale problem, and (v) study computational efficiency of the proposed method (added as a result of the revision).

### 2014-07-11 omrp_lb_rev.tex
Finishing up EinC, AE, Ref1 responses
and a final pass through Ref2 responses and comments -- Still not done.

\item[(ii)] We provide conditions under which the resulting optimality gap estimators exhibit reduced variance when compared to their counterparts obtained via nonoverlapping batches.
We show that the resulting point estimators are strongly consistent and the interval estimators are asymptotically valid. 
These results require new proofs because of the presence of the optimization operator.

@@

\item[(iv)] As the degree of overlap increases, the number of batches and optimization problems solved increases because we need to solve a sampling problem per batch. 
At the same time, variances of the estimators are lowered, and when warm starting is available, solution time per batch can decrease, resulting in efficiency gains. 
We explore the trade-off between computation time and lower variances.
% and efficiency gains. 
%We explore the overall computation time versus lower variances and efficiency gains trade-off. 
We provide guidelines on selecting the degree of overlap based on computational efficiency and demonstrate efficiency on a newsvendor problem. 
\end{itemize}

@@

We wish to apply overlapping batch means to MRP. 
We begin by discussing several differences in this setting compared to the classical simulation setting.  
In simulation output analysis, the point of interest is that of estimating the variance of the sample mean of a covariance stationary process.  
We are interested in estimating the variance of an optimality gap estimator.  
An optimality gap estimator not only has a sample mean $(\frac{1}{m} \sum_{i=1}^m f(\xh,\xiti))$ but also a minimized sample mean $(\zs_m)$.  
Minimization changes the statistical properties of sample means, complicating the asymptotic analysis.  
%For instance, the central limit theorem may not hold for a minimized sample mean, even though it holds for each $\x \in \X$.  
%Therefore, analysis of the asymptotic benefits of overlapping becomes more complicated.
We overcome this technical difficulty by approximating the optimality gap estimators by their nonoptimized counterparts (see \S \ref{subsec:nonO}).  
The nonoptimized counterparts are regular overlapping batch means estimators (without any optimization present); therefore, they have the desired statistical properties of approximately the same bias and lower variances compared to their nonoverlapping versions. 
We establish convergence of OMRP estimators to their nonoptimized counterparts (see \S \ref{subsec:conv}--\ref{ssec:validity}); 
thereby establishing the benefits of overlapping in this context.

>>>>

\item[(ii)] We provide conditions under which the resulting optimality gap estimators exhibit reduced variance when compared to their counterparts obtained via nonoverlapping batches.
We also show that the resulting point estimators are strongly consistent and that the interval estimators are asymptotically valid. 
These results require new proofs because of the presence of the optimization operator.

@@

\item[(iv)] As the degree of overlap increases, the number of batches and optimization problems solved increases because we need to solve a sampling problem per batch. 
At the same time, variances of the estimators are lowered, and when warm starting is available, solution time per batch can decrease, resulting in efficiency gains. 
We explore this trade-off between computation time and lower variances.
% and efficiency gains. 
%We explore the overall computation time versus lower variances and efficiency gains trade-off. 
Our results indicate that when warm-starting is effective, overlapping is computationally more efficient relative to estimators formed via nonoverlapping batches.
%Our results indicate that overlapping can indeed increase computational efficiency, especially when warm-starting is effective. 
We provide guidelines on selecting the degree of overlap based on computational efficiency and demonstrate efficiency on a newsvendor problem. 
\end{itemize}

@@

We wish to apply overlapping batch means to MRP. 
We begin by discussing several differences in this setting compared to the classical simulation setting.  
In simulation output analysis, the point of interest is that of estimating the variance of the sample mean of a covariance stationary process.  
We are interested in estimating the variance of an optimality gap estimator.  
An optimality gap estimator not only has a sample mean $(\frac{1}{m} \sum_{i=1}^m f(\xh,\xiti))$ but also a minimized sample mean; see, e.g., the second term on the right-hand side of (\ref{eq:mrp_gap}).  
Minimization changes the statistical properties of sample means, complicating the asymptotic analysis.  
%For instance, the central limit theorem may not hold for a minimized sample mean, even though it holds for each $\x \in \X$.  
%Therefore, analysis of the asymptotic benefits of overlapping becomes more complicated.
We overcome this technical difficulty by approximating the optimality gap estimators by their nonoptimized counterparts (see \S \ref{subsec:nonO}).  
The nonoptimized counterparts are regular overlapping batch means estimators (without any optimization); therefore, they have the desired statistical properties of approximately the same bias and lower variances compared to their nonoverlapping versions. 
To show that the same asymptotic properties hold for OMRP estimators, we establish convergence of OMRP estimators to their nonoptimized counterparts (see \S \ref{subsec:conv}--\ref{ssec:validity}).
